A Three-Step Approach for Assessing Landscape Connectivity via Simulated Dispersal: African Wild Dog Case Study

### 2022-04-29 Chapter_1/04_Manuscript/Manuscript.tex
upd

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}

@@

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{A Thee-Step Approach for Assessing Landscape Connectivity via Simulated
Dispersal: African Wild Dog Case Study}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Dominik M. Behr\textsuperscript{1,2}
}

@@

\newpage
\begin{abstract}
%------------------------------------------------------------------------------
%  LONG VERSION
%------------------------------------------------------------------------------
\noindent 1. Dispersal of individuals contributes to long-term population
persistence, yet requires a sufficient degree of landscape connectivity. To
date, connectivity has mainly been investigated using least-cost analysis and
circuit theory, two methods that make assumptions that are hardly applicable to
dispersal. Least-cost analysis assumes that animals move towards a known
endpoint and are knowledgeable about the most favorable route. Circuit theory
presumes a complete random walk and fails to incorporate directional
persistence. While these assumptions can be relaxed by explicitly simulating
dispersal trajectories across the landscape, a unified approach for such
simulations is lacking.

\noindent 2. Here, we present a simple three-step approach to simulate dispersal
and to assess connectivity using empirical GPS movement data and a set of
habitat covariates. In step one, we use integrated step-selection functions to
fit a mechanistic movement model describing habitat and movement preferences of
dispersing individuals. In step two, we apply the parameterized model to
simulate dispersal across the study area. In step three, we derive a set of
three complementary connectivity maps; a heatmap highlighting frequently
traversed areas, a betweenness map pinpointing dispersal corridors, and a map of
inter-patch connectivity indicating the presence and intensity of functional
links between habitat patches. We demonstrate the applicability of the proposed
three-step approach in a case study in which we use GPS data collected on
dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting the
Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.

\noindent 3. Using step-selection functions we successfully parametrized a
detailed dispersal model that described dispersing individuals habitat and
movement preferences, as well as potential interactions among the two. The model
substantially outperformed a model that omitted such interactions and enabled us
to simulate 80,000 dispersal trajectories across the study area.

\noindent 4. By explicitly simulating individuals, our approach not only
requires fewer unrealistic assumptions about dispersal, but also permits the
calculation of a variety of connectivity metrics that together provide a
comprehensive view of landscape connectivity. In our case study, the three
derived connectivity maps revealed several wild dog dispersal hotspots and
corridors across the extent of our study area, yet each map highlighted a
different aspect of landscape connectivity, thus emphasizing their complementary
nature. Overall, our case study demonstrates that a simulation-based approach
offers a simple yet powerful alternative to traditional connectivity modeling
techniques. Our approach is therefore useful for a variety of applications in
ecological, evolutionary, and conservation research.

@@

\onehalfspacing
\tableofcontents
\doublespacing

@@

% Importance of Dispersal & Connectivity
\subsection{Importance of Connectivity \& Connectivity Models}
Dispersal of individuals is a vital process that allows species to maintain
genetic diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and to
colonize unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the
ability to disperse depends on a sufficient degree of landscape connectivity
\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
dispersal corridors that promote connectivity a task of fundamental importance
\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. Identifying dispersal corridors
not only necessitates a comprehensive understanding of the factors that limit
dispersal, but also an appropriate model to estimate connectivity
\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
used connectivity models are least-cost path analysis (LCPA;
\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
McRae.2008}). Unfortunately, both models rest on assumptions that appear
unsuitable for dispersers, thus calling for the development of alternative
approaches. One promising alternative is to assess landscape connectivity via
simulated dispersal trajectories generated from individual-based movement models
(IBMMs, \citealp{Diniz.2019}). However, IBMMs require a large number of
subjective modeling decisions, thus making among-system comparisons difficult.

% Issues with Both Methods
\subsection{Issues with Traditional Connectivity Models}
Traditional connectivity models make assumptions that are rarely met for
dispersers. LCPA, for instance, assumes that individuals move towards a
preconceived endpoint and choose a cost-minimizing route accordingly
\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
migrating animals, it is unlikely to hold for dispersers, as dispersers
typically move across unfamiliar territory towards an unknown endpoint
\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
according to a random walk, entailing that autocorrelation between subsequent
movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
autocorrelated movements are regularly observed \citep{Cozzi.2020,
Hofmann.2021}, meaning that dispersal trajectories are usually strongly
directional. Moreover, because both models require static permeability or
resistance surfaces as input, they are unable to reflect the temporal dimension
of dispersal and therefore fail to allow statements about the expected duration
for moving between habitat patches \citep{Martensen.2017, Diniz.2019}.

% What about IBMMs?
\subsection{What about IBMMs?}
The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
using IBMMs and by converting simulated trajectories into meaningful measures of
connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
explicitly simulate how individuals move across and interact with the
encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
between movement behavior and habitat conditions \citep{Avgar.2016}. This
strictly shifts the focus from a structural to a more functional view on
connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
Furthermore, IBMMs generate movement sequentially, i.e. they generate a series
of steps, so that the temporal dimension of dispersal becomes explicit and
allows modeling autocorrelation between successive steps \citep{Diniz.2019}.
Finally, simulations from IBMMs do not enforce movement or connections towards
preconceived endpoints but allow individuals to adjust their route  ``on the
go'', thereby preventing biases arising from misplaced endpoints. Despite these
advantages, a unifying approach to simulate dispersal and assess connectivity
using IBMMs is lacking. Considering the large number of subjective decisions
entailed by IBMMs, an approach that streamlines and standardizes the application
of dispersal simulations to assess connectivity will, however, be critical to
safeguard comparability among studies.

% Proposed Solution: Three-Step approach
\subsection{Proposed Solution: Three-Step Approach}
Here, we propose and exemplify a simple three-step approach for simulating
dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
step one, we combine GPS movement data of dispersing individuals with habitat
covariates to fit a mechanistic movement model via integrated step-selection
functions (ISSFs, \citealp{Avgar.2016}). We chose to use ISSFs because the
framework not only allows inference on the study species' habitat kernel (i.e.
its habitat preferences), but also its movement kernel (i.e. its movement
preferences/capabilities) and potential interactions among the two
\citep{Avgar.2016, Fieberg.2021}. In step two, we use the parametrized movement
model to simulate dispersal across the study area. Comparable simulations have
already been applied to estimate steady-state utilization distributions of
resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
connectivity, yet disregarding interdependencies between habitat and movement
kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert the
simulated trajectories into three complementary connectivity maps; (i) a heatmap
revealing frequently traversed areas (e.g. \citealp{Hauenstein.2019,
Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
inter-patch connectivity, depicting the presence and intensity of functional
links between habitat patches, as well as the average dispersal duration
required to realize those connections (e.g. \citealp{Gustafson.1996,
Kanagaraj.2013}).

@@

% Introduction of the study species
\subsection{Case Study}
We showcase the application of the proposed approach using GPS movement data
collected on dispersing African wild dogs (\textit{Lycaon pictus}). The African
wild dog is a highly mobile species whose population persistence heavily relies
on the availability of large, natural or semi-natural landscapes and a
sufficient degree of connectivity among remaining subpopulations. Once common
throughout sub-Saharan Africa, this species has disappeared from much of its
historic range, largely due to human persecution, habitat fragmentation, and
disease outbreaks \citep{Woodroffe.2012}. Wild dogs typically disperse in
single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
dispersing several hundred kilometers \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}. Although previous studies have investigated connectivity for this
species using LCPA \citep{Hofmann.2021} and CT \citep{Brennan.2020}, a more
comprehensive and mechanistic understanding of dispersal and connectivity is
missing (but see \citealp{Creel.2020}). Nevertheless, with about 6,000
free-ranging wild dogs remaining in fragmented subpopulations
\citep{Woodroffe.2012}, reliable information on dispersal behavior and landscape
connectivity is essential for the conservation of this endangered carnivore. We
anticipated that a connectivity assessment based upon our three-step approach
would overcome several of the conceptual shortcomings of traditional
connectivity models, while providing a more detailed view on movement behavior
during dispersal its implications for landscape connectivity.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_GraphicalAbstract2.pdf}
    \caption{Flowchart of the simulation-based connectivity analysis as proposed
    in this article. First, GPS data and habitat covariates must be
    collected. The combined data is then analyzed in an integrated step
    selection model, which enables the parametrization of the focal species'
    habitat and movement kernels and results in a mechanistic movement model.
    The parametrized model is then treated as an individual-based movement model
    and used to simulate dispersal trajectories. Ultimately, simulated
    trajectories serve to produce a set of maps that are pertinent to landscape
    connectivity. This includes a heatmap, indicating the traversal frequency
    across each spatial unit of the study area, a betweenness map, highlighting
    movement corridors and bottlenecks, and, finally, an inter-patch
    connectivity map, where the frequency of connections and their average
    duration can be depicted.}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.png}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
        \end{scope}
    \end{tikzpicture}
    \caption{Illustration of the study area located in southern Africa. (a) The
    study area was confined by a bounding box spanning the entire KAZA-TFCA
    which comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia.
    Data on remaining wild dog populations has been sourced from
    \cite{Woodroffe.2012}. (b) The KAZA-TFCA currently represents the world's
    largest terrestrial transfrontier conservation area, covering a total area
    of 520'000 km\textsuperscript{2}. Its main purpose is to re-establish
    connectivity between already-existing NPs (dark green) and other protected
    areas (light green).}
    \label{StudyArea}
  \end{center}
\end{figure}

@@

We performed all data preparations, spatial computations, and statistical
analysis in {\tt R}, version 3.6.6 \citep{R.2020}. Some helper functions were
written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
\citep{Eddelbuettel.2011, Eddelbuettel.2013}.

@@

Along each realized and random step, we extracted values from underlying habitat
covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
computed averages of each covariate along the steps. Besides extracting
\textit{habitat covariates}, we also computed movement metrics that we used as
\textit{movement covariates} in the ISSF models \citep{Avgar.2016,
Fieberg.2021}. Specifically, we computed the step length (\textsf{sl}), its
natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
angle (\textsf{cos(ta)}), which is a measure of directionality
\citep{Turchin.1998}, for each step. Because wild dog activity is low during the
hot midday hours \citep{Cozzi.2012}, we additionally created the variable
\textsf{LowActivity}, indicating whether a step was realized during periods of
low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
(17:00 to 09:00 local time). To facilitate model convergence, we standardized
all continuous covariates to a mean of zero and a standard deviation of one.
Correlations among covariates were low (\(|r| < 0.6\); \citealp{Latham.2011}),
so we retained all of them for modeling.

@@

\subsection{Step 3 - Connectivity Maps}
\subsubsection{Heatmap}
To identify dispersal hotspots within the study area, we created a heatmap
indicating the absolute frequency at which different areas were traversed by
simulated dispersal trajectories (e.g. \citealp{Peer.2008, Hauenstein.2019,
Zeller.2020}). Specifically, we rasterized all simulated trajectories onto a
raster with 1 km x 1 km resolution and tallied resulting layers into a single
map. This procedure ensured that every trajectory was only counted once, even if
it traversed the same raster-cell multiple times, thus reducing potential biases
caused by individuals that were surrounded by unfavorable habitat and ``moved in
circles''. To achieve high performance rasterization, we used the R-package {\tt
terra} \citep{Hijmans.2021b}.

@@

\section{Results}
\subsection{Movement Model}
The most parsimonious movement model consisted of movement covariates, habitat
covariates, as well as several of their interactions, thus suggesting that
movement behavior during dispersal depended on habitat conditions
(\Cref{MovementModel}a, Table S1 and Table S2). Although multiple models
received an AIC weight \(>\) 0 (Table S1), we only considered results from the
most parsimonious model for simplicity. This decision only marginally influenced
subsequent steps as all models with positive AIC weights retained similar
covariates (Table S1). The k-fold cross-validation showed that the final model
substantially outperformed a random guess and provided reliable predictions
(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
random}\) did not overlap). Moreover, the model correctly assigned high
selection scores to realized steps (\Cref{MovementModel}b), indicating a good
fit between predictions and observations. As can be taken from the Spearman rank
correlation coefficient, the inclusion of several interactions between movement
and habitat covariates significantly improved model performance (\(\bar{r}_{s,
realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)), compared to the base model
(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
\citealp{Hofmann.2021})

@@

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmap.png}
  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
  integrated step-selection model that we fitted to the movement data of
  dispersing African wild dogs. To generate the heatmap, we rasterized and
  tallied all simulated trajectories. Consequently, the map highlights areas
  that are frequently traversed by virtual dispersers. For spatial reference we
  plotted a few selected NPs (dark gray). Additional heatmaps showing the
  traversal frequency when individuals move fewer than 2'000 steps are provided
  in Figure S4.}
  \label{Heatmap}
\end{figure}

>>>>

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

@@

% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{A Three-Step Approach for Assessing Landscape Connectivity via Simulated
Dispersal: African Wild Dog Case Study}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
}

@@

\newpage
\begin{abstract}
%------------------------------------------------------------------------------
%  LONG VERSION
%------------------------------------------------------------------------------
\noindent \textbf{Context} \quad Dispersal of individuals contributes to
long-term population persistence, yet requires a sufficient degree of landscape
connectivity. To date, connectivity has mainly been investigated using
least-cost analysis and circuit theory, two methods that make assumptions that
are hardly applicable to dispersal. While these assumptions can be relaxed by
explicitly simulating dispersal trajectories across the landscape, a unified
approach for such simulations is lacking.

\noindent \textbf{Objectives} \quad Here, we propose and apply a simple
three-step approach to simulate dispersal and to assess connectivity using
empirical GPS movement data and a set of habitat covariates.

\noindent \textbf{Methods} \quad In step one of the proposed approach, we use
integrated step-selection functions to fit a mechanistic movement model
describing habitat and movement preferences of dispersing individuals. In step
two, we apply the parameterized model to simulate dispersal across the study
area. In step three, we derive three complementary connectivity maps; a heatmap
highlighting frequently traversed areas, a betweenness map pinpointing dispersal
corridors, and a map of inter-patch connectivity indicating the presence and
intensity of functional links between habitat patches. We demonstrate the
applicability of the proposed three-step approach in a case study in which we
use GPS data collected on dispersing African wild dogs (\textit{Lycaon pictus})
inhabiting northern Botswana.

\noindent \textbf{Results} \quad Using step-selection functions we successfully
parametrized a detailed dispersal model that described dispersing individuals'
habitat and movement preferences, as well as potential interactions among the
two. The model substantially outperformed a model that omitted such interactions
and enabled us to simulate 80,000 dispersal trajectories across the study area.

\noindent \textbf{Conclusion} \quad By explicitly simulating dispersal
trajectories, our approach not only requires fewer unrealistic assumptions about
dispersal, but also permits the calculation of multiple connectivity metrics
that together provide a comprehensive view of landscape connectivity. In our
case study, the three derived connectivity maps revealed several wild dog
dispersal hotspots and corridors across the extent of our study area. Each map
highlighted a different aspect of landscape connectivity, thus emphasizing their
complementary nature. Overall, our case study demonstrates that a
simulation-based approach offers a simple yet powerful alternative to
traditional connectivity modeling techniques. It is therefore useful for a
variety of applications in ecological, evolutionary, and conservation research.

@@

% \onehalfspacing
% \tableofcontents
\doublespacing

@@

% Importance of Dispersal & Connectivity
% \subsection{Importance of Connectivity \& Connectivity Models}
Dispersal of individuals is a vital process that allows species to maintain
genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013},
rescue non-viable populations \citep{Brown.1977}, and to colonize unoccupied
habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability to disperse
depends on a sufficient degree of landscape connectivity \citep{Fahrig.2003,
Clobert.2012}, making the identification and protection of dispersal corridors
that promote connectivity a task of fundamental importance \citep{Doerr.2011,
Rudnick.2012}. Identifying dispersal corridors not only necessitates a
comprehensive understanding of the factors that limit dispersal, but also an
appropriate model to estimate connectivity \citep{Baguette.2013, Vasudev.2015,
Hofmann.2021}. To date, the most commonly used connectivity models are
least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory
(CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest on
assumptions that appear unsuitable for dispersers, thus calling for the
development of alternative approaches. One promising alternative is to assess
landscape connectivity via simulated dispersal trajectories generated from
individual-based movement models (IBMMs, \citealp{Diniz.2019}). However, IBMMs
require a large number of subjective modeling decisions, thus making
among-system comparisons difficult.

% Issues with Both Methods
% \subsection{Issues with Traditional Connectivity Models}
Traditional connectivity models make assumptions that are rarely met for
dispersers. LCPA, for instance, assumes that individuals move towards a
preconceived endpoint and choose a cost-minimizing route accordingly
\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
migrating animals, it is unlikely to hold for dispersers, as dispersers
typically move across unfamiliar territory towards an unknown endpoint
\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
according to a random walk, entailing that autocorrelation between subsequent
movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
autocorrelated movements are regularly observed \citep{Cozzi.2020,
Hofmann.2021}, meaning that dispersal trajectories are usually strongly
directional. Moreover, because both models require static permeability or
resistance surfaces as input, they are unable to reflect the temporal dimension
of dispersal and therefore fail to allow statements about the expected duration
for moving between habitat patches \citep{Martensen.2017, Diniz.2019}.

% What about IBMMs?
% \subsection{What about IBMMs?}
The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
using IBMMs and by converting simulated trajectories into meaningful measures of
connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
explicitly simulate how individuals move across and interact with the
encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
between movement behavior and habitat conditions \citep{Avgar.2016}. This
strictly shifts the focus from a structural to a more functional view on
connectivity \citep{Tischendorf.2000}. Furthermore, IBMMs generate movement
sequentially, i.e. they generate a series of steps, so that the temporal
dimension of dispersal becomes explicit and allows modeling autocorrelation
between successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do
not enforce movement or connections towards preconceived endpoints but allow
individuals to adjust their route  ``on the go'', thereby preventing biases
arising from misplaced endpoints. Despite these advantages, a unifying approach
to simulate dispersal and assess connectivity using IBMMs is lacking.
Considering the large number of subjective decisions entailed by IBMMs, an
approach that streamlines and standardizes the application of dispersal
simulations to assess connectivity will, however, be critical to safeguard
comparability among studies.

% Proposed Solution: Three-Step approach
% \subsection{Proposed Solution: Three-Step Approach}
Here, we propose and exemplify a simple three-step approach for simulating
dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
step one, we combine GPS movement data of dispersing individuals with habitat
covariates to fit a mechanistic movement model via integrated step-selection
functions (ISSFs, \citealp{Avgar.2016}). We chose to use ISSFs because the
framework not only allows inference on the study species' habitat kernel (i.e.
its habitat preferences), but also its movement kernel (i.e. its movement
preferences/capabilities) and potential interactions among the two
\citep{Avgar.2016, Fieberg.2021}. In step two, we use the parametrized movement
model to simulate dispersal across the study area. Comparable simulations have
already been applied to estimate steady-state utilization distributions of
resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
connectivity, yet disregarding interdependencies between habitat and movement
kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert the
simulated trajectories into three complementary connectivity maps; (i) a heatmap
revealing frequently traversed areas (e.g. \citealp{Hauenstein.2019,
Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
inter-patch connectivity, depicting the presence and intensity of functional
links between habitat patches, as well as the average dispersal duration
required to realize those connections (e.g. \citealp{Gustafson.1996,
Kanagaraj.2013}).

@@

% Introduction of the study species
% \subsection{Case Study}
We showcase the application of the proposed approach using GPS movement data
collected on dispersing African wild dogs (\textit{Lycaon pictus}). The African
wild dog is a highly mobile species whose population persistence heavily relies
on the availability of large, natural or semi-natural landscapes and a
sufficient degree of connectivity among remaining subpopulations. Once common
throughout sub-Saharan Africa, this species has disappeared from much of its
historic range, largely due to human persecution, habitat fragmentation, and
disease outbreaks \citep{Woodroffe.2012}. Wild dogs typically disperse in
single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
dispersing several hundred kilometers \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}. Although previous studies have investigated connectivity for this
species using LCPA \citep{Hofmann.2021} and CT \citep{Brennan.2020}, a more
comprehensive and mechanistic understanding of dispersal and connectivity is
missing (but see \citealp{Creel.2020}). Nevertheless, with about 6,000
free-ranging wild dogs remaining in fragmented subpopulations
\citep{Woodroffe.2012}, reliable information on dispersal behavior and landscape
connectivity is essential for the conservation of this endangered carnivore. We
anticipated that a connectivity assessment based upon our three-step approach
would overcome several of the conceptual shortcomings of traditional
connectivity models, while providing a more detailed view on movement behavior
during dispersal its implications for landscape connectivity.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
    data and habitat covariates must be collected. The combined data is then
    analyzed using an integrated step selection model. The parametrized model is
    then treated as an individual-based movement model and used to simulate
    dispersal trajectories. Ultimately, simulated trajectories are translated
    into a set of maps that are pertinent to landscape connectivity. This
    includes a heatmap, indicating the traversal frequency across each spatial
    unit of the study area, a betweenness map, highlighting movement corridors
    and bottlenecks, and, finally, an inter-patch connectivity map, where the
    frequency of connections and their average duration can be depicted.}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.png}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
        \end{scope}
    \end{tikzpicture}
    \caption{Illustration of the study area in southern Africa. (a) The study
    area was confined by a bounding box spanning the entire KAZA-TFCA which
    comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. Data on
    remaining wild dog populations (orange) has been sourced from
    \cite{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
    terrestrial transfrontier conservation area and covers a total area of
    520'000 km\textsuperscript{2}. Its main purpose is to re-establish
    connectivity between already-existing NPs (dark green) and other protected
    areas (light green).}
    \label{StudyArea}
  \end{center}
\end{figure}

@@

We performed all data preparations, spatial computations, and statistical
analysis in {\tt R}, version 3.6.6 \citep{R.2020}. Some helper functions were
written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
\citep{Eddelbuettel.2011, Eddelbuettel.2013, Eddelbuettel.2018}.

@@

Along each realized and random step, we extracted values from underlying habitat
covariate layers and we computed averages of each covariate along the steps.
Besides extracting \textit{habitat covariates}, we also computed movement
metrics that we used as \textit{movement covariates} in the ISSF models
\citep{Avgar.2016, Fieberg.2021}. Specifically, we computed the step length
(\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
relative turning angle (\textsf{cos(ta)}), which is a measure of directionality
\citep{Turchin.1998}, for each step. Because wild dog activity is low during the
hot midday hours \citep{Cozzi.2012}, we additionally created the variable
\textsf{LowActivity}, indicating whether a step was realized during periods of
low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
(17:00 to 09:00 local time). To facilitate model convergence, we standardized
all continuous covariates to a mean of zero and a standard deviation of one.
Correlations among covariates were low (\(|r| < 0.6\); \citealp{Latham.2011}),
so we retained all of them for modeling.

@@

\subsection{Step 3 - Connectivity Maps}
\subsubsection{Heatmap}
To identify dispersal hotspots within the study area, we created a heatmap
indicating the absolute frequency at which different areas were traversed by
simulated dispersal trajectories (e.g. \citealp{Hauenstein.2019, Zeller.2020}).
Specifically, we rasterized all simulated trajectories onto a raster with 1 km x
1 km resolution and tallied resulting layers into a single map. This procedure
ensured that every trajectory was only counted once, even if it traversed the
same raster-cell multiple times, thus reducing potential biases caused by
individuals that were surrounded by unfavorable habitat and ``moved in
circles''. To achieve high performance rasterization, we used the R-package {\tt
terra} \citep{Hijmans.2021b}.

@@

\section{Results}
% \subsection{Movement Model}
The most parsimonious movement model consisted of movement covariates, habitat
covariates, as well as several of their interactions, thus suggesting that
movement behavior during dispersal depended on habitat conditions
(\Cref{MovementModel}a, Table S1 and Table S2). Although multiple models
received an AIC weight \(>\) 0 (Table S1), we only considered results from the
most parsimonious model for simplicity. This decision only marginally influenced
subsequent steps as all models with positive AIC weights retained similar
covariates (Table S1). The k-fold cross-validation showed that the final model
substantially outperformed a random guess and provided reliable predictions
(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
random}\) did not overlap). Moreover, the model correctly assigned high
selection scores to realized steps (\Cref{MovementModel}b), indicating a good
fit between predictions and observations. As can be taken from the Spearman rank
correlation coefficient, the inclusion of several interactions between movement
and habitat covariates significantly improved model performance (\(\bar{r}_{s,
realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)), compared to the base model
(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
\citealp{Hofmann.2021})

@@

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmap.png}
  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
  integrated step-selection model that we fitted to the movement data of
  dispersing African wild dogs. To generate the heatmap, we rasterized and
  tallied all simulated trajectories. Consequently, the map highlights areas
  that are frequently traversed. For spatial reference we plotted a few selected
  NPs (dark gray). Additional heatmaps showing the traversal frequency when
  individuals move fewer than 2'000 steps are provided in Figure S4.}
  \label{Heatmap}
\end{figure}
